---
title: "ACM_3"
output: html_document
date: "2024-03-22"
---

# Assignment 3: Multilevel modelling

```{r setup}
pacman::p_load(tidyverse, cmdstanr, brms)
```

## Introduction

**Experiment description**
The data comes from the social conformity experiment where a participant is asked to rate the trustworthiness of a face. Immediately after having given a rating, the participant is then shown the ratings of the same face by a group of people. After an hour, the participant is asked to rate the face again. The experiment is repeated for a number of faces. The data is collected from cognitive science students during the pandemic.

```{r data}
# Data from cogsci students during the pandemic
df <- read_csv("data/sc_df_clean.csv")

# fixing error in the data (some trials does not get feedback)
```

**Data description**
The data comes from the social conformity experiment (https://pubmed.ncbi.nlm.nih.gov/30700729/), where cogsci students combine their own intuition of trustworthiness of given faces to social information.

The important variables in the data are the following:

| Variable | Description |
|----------|----------|
| FaceID        | Number representing faces [INT]     |
| FirstRating   | Initial trustworthiness rating given by participant [INT]    |
| GroupRating   | Mean trustworthiness rating given by group [INT]    |
| SecondRating  | Second trustworthiness rating given by participant [INT]    |
| Feedback      | Difference between first rating and group rating [INT]    |
| Change        | Difference between first and second rating by participant [INT]    |

**Extra OBS**
Some trials does not get feedback (NAs) - they are a group in themselves.
Some of the feedbacks are 0s - they needs to be fixed.

Possible effect and limitations of the model:
If you rate someone as being 5, and group does the same, you are more likely to rate the person as 5 or higher the next time, because a majority of people rate x person as being above average trustworthy.
The choices of trustworthiness are not independent of each other, but are influenced by the group rating. 

# Two agents: Simple and weighted Bayes
```{r agents}
# Function to make sure the denominator is never 0 (which produces NaN since you can't divide by 0.
logit_scaled <- function(p) {
  log(p / (1 - p + 1e-9))
}

simpleBayes_f <- function(bias, Source1, Source2){
    outcome <- inv_logit_scaled(bias + logit_scaled(Source1 / 8) + logit_scaled(Source2 / 8)) # dividing by 8 to scale the ratings to 0-1 (probability scale)
    return(outcome)
}

WeightedBayes_f <- function(bias, Source1, Source2, w1, w2){
    w1 <- (w1 - 0.5)*2
    w2 <- (w2 - 0.5)*2
    
    # Define or load the inv_logit_scaled and logit_scaled functions here
    outcome <- inv_logit_scaled(bias + w1 * logit_scaled(Source1 / 8) + w2 * logit_scaled(Source2 / 8)) # 
    return(outcome)
}
```

# Creating simple data
```{r simulating simple data}
trials <- 150
participants <- 20
ratings <- 1:8 # the participants can rate the faces from 1 to 8
w1 <- .5 # the weight for own rating
w2 <- .7 # the weight for group rating
bias <- 0 # assuming no bias
feedback <- c(-3, -2 ,0 ,2, 3) # the group rating is randomly chosen from this vector and subtracted from the first rating

sim_data <- tibble()

for (participant in seq(participants)){ # for each participant
    for (i in seq(trials)) { # for each trial
        sim_data <- rbind(sim_data, tibble(
            participant = participant,
            FaceID = i,
            FirstRating = sample(ratings, 1, prob = c(0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.7, 0.4)),
            GroupRating = pmax(1, pmin(8, FirstRating + (sample(feedback, 1)))), # pmin / pmax ensures that the group rating is within 1-8
            # add feedback saying if how much higher or lower the group rating is compared to the first rating
            feedback = GroupRating - FirstRating      
        ))
    }
}
sim_data$w1 <- w1
sim_data$w2 <- w2
sim_data$bias <- bias
```

# Simulating data from 1) simple bayes model
```{r simulate data from simple bayes model}
# Simple Bayes second rating (after having seen the group rating)
for (participant in seq(participants)){ # for each participant
    for (i in seq(trials)) { # for each trial
        sim_data <- sim_data %>% 
            mutate(SecondRating_belief_SB = simpleBayes_f(bias, FirstRating, GroupRating)) %>% # SB = simple bayes
            mutate(SB_binary_rating = ifelse(SecondRating_belief_SB > 0.5, 1, 0)) %>% # binary rating (either trustworthiness or not) Alternativ: rbinom(1,1,SB_binary_rating)
            mutate(SB_number_rating = round(SecondRating_belief_SB * 8)) # number rating (1-8)
    }
}
```

# Simulating data from 2) weighted bayes model
```{r simulating data from weighted bayes model}
# Weighted Bayes second rating (after having seen the group rating)
for (participant in seq(participants)){ # for each participant
    for (i in seq(trials)) { # for each trial
        sim_data <- sim_data %>% 
            mutate(SecondRating_belief_WB = WeightedBayes_f(bias, FirstRating, GroupRating, w1, w2)) %>% # WB = weighted bayes
            mutate(WB_binary_rating = ifelse(SecondRating_belief_WB > 0.5, 1, 0)) %>% # binary rating (either trustworthiness or not) Alternativ: rbinom(1,1,WB_binary_rating)
            mutate(WB_number_rating = round(SecondRating_belief_WB * 8)) # number rating (1-8)
    }
}
```

# plotting the second choices for both models
```{r}
pacman::p_load(gridExtra)

plt1 <- ggplot(sim_data, aes(SecondRating_belief_SB)) +
    geom_histogram(bins = 10, alpha = 0.2, color = "black") +
    labs(title = "Simple Bayes model",
         x = "Second rating",
         y = "Frequency")
    theme_bw()

plt2 <- ggplot(sim_data, aes(SecondRating_belief_WB)) +
    geom_histogram(bins = 10, alpha = 0.2, color = "black") +
    theme_bw() +
    labs(title = "Simple Bayes model",
         x = "Second rating",
         y = "Frequency")

pl3 <- ggplot(sim_data, aes(FirstRating, SecondRating_belief_SB, color = GroupRating, group = GroupRating)) +
    geom_line() + 
    labs(title = "Simple Bayes model",
         x = "First rating",
         y = "Second rating") +
    theme_bw()

plt4 <- ggplot(sim_data, aes(FirstRating, SecondRating_belief_WB, color = GroupRating, group = GroupRating)) +
    geom_line() + 
    labs(title = "Weighted Bayes model",
         x = "First rating",
         y = "Second rating") +
    theme_bw()

grid.arrange(plt1, plt2, pl3, plt4, ncol = 2, nrow = 2)
plt5 <- grid.arrange(plt1, plt2, pl3, plt4, ncol = 2, nrow = 2)


# save to drive
ggsave("figures/SecondChoice_Sim_Data.png", plt5, width = 10, height = 10)
```

Question:
Why are the lines not going from 0-8?
What is wrong with the WB model plot?

# STAN modelling of SB model
```{r}
file <- file.path("stan_models/SB_model.stan")

SB_mod <- cmdstan_model(file, cpp_options = list(stan_threads = TRUE), stanc_options = list("O1"), pedantic = TRUE)
```

# Applying the SB model to the simulated data

Preparing data
```{r preparing data}
stan_data <- list(
    N = nrow(sim_data),
    Source1 = sim_data$FirstRating/8, # dividing by 8 to scale the ratings to 0-1 (probability scale)
    Source2 = sim_data$GroupRating/8, # dividing by 8 to scale the ratings to 0-1 (probability scale)
    w1 = sim_data$w1,
    w2 = sim_data$w2,
    bias = sim_data$bias,
    y = sim_data$SB_binary_rating
)
```

```{r}
view(sim_data)
```

Fitting SB model
```{r}
SB_fit <- SB_mod$sample(
    data = stan_data,
    seed = 123,
    chains = 2,
    parallel_chains = 2,
    threads_per_chain = 2,
    iter_warmup = 2000,
    iter_sampling = 2000,
    refresh = 1000,
    max_treedepth = 20, 
    adapt_delta = 0.99 
    )
```

Assessing SB fit 
```{r}
SB_fit$summary()

SB_fit$cmdstan_diagnose()

SB_fit$loo()
```

Plotting SB fit 
```{r}
# Extracting draws
draws_SB <- as_draws_df(SB_fit$draws())

# plotting markov chains for bias
plt1 <- ggplot(draws_SB, aes(.iteration, bias, group = .chain, color = .chain)) +
  geom_line(alpha = 0.5) +
  labs(title = "Markov chains for bias",
       x = "Iteration",
       y = "Bias") +
  theme_classic()

# plotting density of bias, prior and posterior
plt2 <- ggplot(draws_SB) +
  geom_density(aes(bias, fill = "Posterior"), alpha = 0.9) +
  geom_density(aes(bias_prior, fill = "Prior"), alpha = 0.9) +
  geom_vline(xintercept = sim_data$bias[1]) + # true value of bias
  theme_bw() +
  labs(title = "Bias posterior and prior",
       x = "Bias",
       y = "Density")

plt3 <- grid.arrange(plt1, plt2, nrow = 2)
plt3

# save to drive
ggsave("figures/MarkovChains_and_density_SB.png", plt3, width = 10, height = 10)
```

# Applying the WB model to the simulated data
```{r}
file <- file.path("stan_models/WB_model.stan")

WB_mod <- cmdstan_model(file, cpp_options = list(stan_threads = TRUE), stanc_options = list("O1"), pedantic = TRUE)
```

Fitting model
```{r}
WB_fit <- WB_mod$sample(
    data = stan_data,
    seed = 123,
    chains = 2,
    parallel_chains = 2,
    threads_per_chain = 2,
    iter_warmup = 2000,
    iter_sampling = 2000,
    refresh = 1000,
    max_treedepth = 20, 
    adapt_delta = 0.99 
    )
```

Assessing WB fit
```{r}
WB_fit$summary()

WB_fit$cmdstan_diagnose()

WB_fit$loo()
```

Plotting WB fit
```{r}
# Extracting draws
draws_WB <- as_draws_df(WB_fit$draws())

# plotting markov chains for bias
plt1 <- ggplot(draws_WB, aes(.iteration, bias, group = .chain, color = .chain)) +
  geom_line(alpha = 0.5) +
  labs(title = "Markov chains for bias",
       x = "Iteration",
       y = "Bias") +
  theme_classic()

plt2 <- ggplot(draws_WB, aes(.iteration, w1, group = .chain, color = .chain)) +
    geom_line(alpha = 0.5) +
    labs(title = "Markov chains for w1",
         x = "Iteration",
         y = "w1") +
    theme_classic()

plt3 <- ggplot(draws_WB, aes(.iteration, w2, group = .chain, color = .chain)) +
    geom_line(alpha = 0.5) +
    labs(title = "Markov chains for w2",
         x = "Iteration",
         y = "w2") +
    theme_classic()

plt4 <- grid.arrange(plt1, plt2, plt3, nrow = 3)
plt4

# save to drive
ggsave("figures/MarkovChains_WB.png", plt4, width = 10, height = 10)
```

Plotting density of bias, w1, w2: prior and posterior
```{r}
plt1 <- ggplot(draws_WB) +
  geom_density(aes(bias, fill = "Posterior"), alpha = 0.9) +
  geom_density(aes(bias_prior, fill = "Prior"), alpha = 0.9) +
  geom_vline(xintercept = sim_data$bias[1]) + # true value of bias
  theme_bw() +
  labs(title = "Bias posterior and prior",
       x = "Bias",
       y = "Density")

plt2 <- ggplot(draws_WB) +
    geom_density(aes(w1, fill = "Posterior"), alpha = 0.9) +
    geom_density(aes(w1_prior, fill = "Prior"), alpha = 0.9) +
    geom_vline(xintercept = sim_data$w1[1]) + # true value of bias
    theme_bw() +
    labs(title = "w1 posterior and prior",
         x = "w1",
         y = "Density")

plt3 <- ggplot(draws_WB) +
    geom_density(aes(w2, fill = "Posterior"), alpha = 0.9) +
    geom_density(aes(w2_prior, fill = "Prior"), alpha = 0.9) +
    geom_vline(xintercept = sim_data$w2[1]) + # true value of bias
    theme_bw() +
    labs(title = "w2 posterior and prior",
         x = "w2",
         y = "Density")
        
plt4 <- grid.arrange(plt1, plt2, plt3, nrow = 3)
plt4

# save to drive
ggsave("figures/Density_WB.png", plt4, width = 10, height = 10)
```

Question: 
Why is the posterior for bias so off???? Despite the fact that the markov chain looks alright...?


# Fitting model on actual cogsci data wuhu

Preparing data
```{r}
df$Second_rating_binary <- ifelse(df$SecondRating > 4, 1, 0) # binary rating (either trustworthiness or not)
df <- df %>% filter(GroupRating != 0) # cleaning up data (jvf. Riccardos comment on class)

stan_data_real <- list(
    N = nrow(df),
    Source1 = pmin(pmax(df$FirstRating / 8, 0.01), 0.99), # Avoiding exact 0/1
    Source2 = pmin(pmax(df$GroupRating / 8, 0.01), 0.99), # Avoiding exact 0/1
    y = df$Second_rating_binary
)

```

Fitting to SB model
```{r}
SB_fit <- SB_mod$sample(
    data = stan_data_real,
    seed = 123,
    chains = 2,
    parallel_chains = 2,
    threads_per_chain = 2,
    iter_warmup = 2000,
    iter_sampling = 2000,
    refresh = 1000,
    max_treedepth = 20, 
    adapt_delta = 0.99 
    )
```

Assessing SB fit
```{r}
SB_fit$summary()

SB_fit$cmdstan_diagnose()

SB_fit$loo()
```

Plotting SB fit
```{r}
# Extracting draws
draws_SB_real <- as_draws_df(SB_fit$draws())

# plotting markov chains for bias
plt1 <- ggplot(draws_SB_real, aes(.iteration, bias, group = .chain, color = .chain)) +
  geom_line(alpha = 0.5) +
  labs(title = "Markov chains for bias",
       x = "Iteration",
       y = "Bias") +
  theme_classic()

# plotting density of bias, prior and posterior
plt2 <- ggplot(draws_SB_real) +
  geom_density(aes(bias, fill = "Posterior"), alpha = 0.9) +
  geom_density(aes(bias_prior, fill = "Prior"), alpha = 0.9) +
  geom_vline(xintercept = 0) + # true value of bias
  theme_bw() +
  labs(title = "Bias posterior and prior",
       x = "Bias",
       y = "Density")

plt3 <- grid.arrange(plt1, plt2, nrow = 2)
plt3

# save to drive
ggsave("figures/MarkovChains_and_density_SB_real.png", plt3, width = 10, height = 10)
```

Fitting to WB model
```{r}
WB_fit_real <- WB_mod$sample(
    data = stan_data_real,
    seed = 123,
    chains = 2,
    parallel_chains = 2,
    threads_per_chain = 2,
    iter_warmup = 2000,
    iter_sampling = 2000,
    refresh = 1000,
    max_treedepth = 20, 
    adapt_delta = 0.99 
    )
```

Assessing WB fit
```{r}
WB_fit_real$summary()

WB_fit_real$cmdstan_diagnose()

WB_fit_real$loo()
```

Plotting WB fit
```{r}
# Extracting draws
draws_WB_real <- as_draws_df(WB_fit_real$draws())

# plotting markov chains for bias
plt1 <- ggplot(draws_WB_real, aes(.iteration, bias, group = .chain, color = .chain)) +
  geom_line(alpha = 0.5) +
  labs(title = "Markov chains for bias",
       x = "Iteration",
       y = "Bias") +
  theme_classic()

plt2 <- ggplot(draws_WB_real, aes(.iteration, w1, group = .chain, color = .chain)) +
    geom_line(alpha = 0.5) +
    labs(title = "Markov chains for w1",
         x = "Iteration",
         y = "w1") +
    theme_classic()

plt3 <- ggplot(draws_WB_real, aes(.iteration, w2, group = .chain, color = .chain)) +
    geom_line(alpha = 0.5) +
    labs(title = "Markov chains for w2",
         x = "Iteration",
         y = "w2") +
    theme_classic()

plt4 <- grid.arrange(plt1, plt2, plt3, nrow = 3)
plt4

# save to drive
ggsave("figures/MarkovChains_WB_real.png", plt4, width = 10, height = 10)

```

Plotting density of bias, w1, w2: prior and posterior
```{r}
plt1 <- ggplot(draws_WB_real) +
  geom_density(aes(bias, fill = "Posterior"), alpha = 0.9) +
  geom_density(aes(bias_prior, fill = "Prior"), alpha = 0.9) +
  geom_vline(xintercept = 0) + # true value of bias
  theme_bw() +
  labs(title = "Bias posterior and prior",
       x = "Bias",
       y = "Density")

plt2 <- ggplot(draws_WB_real) +
    geom_density(aes(w1, fill = "Posterior"), alpha = 0.9) +
    geom_density(aes(w1_prior, fill = "Prior"), alpha = 0.9) +
    geom_vline(xintercept = 0.5) + # true value of bias
    theme_bw() +
    labs(title = "w1 posterior and prior",
         x = "w1",
         y = "Density")

plt3 <- ggplot(draws_WB_real) +
    geom_density(aes(w2, fill = "Posterior"), alpha = 0.9) +
    geom_density(aes(w2_prior, fill = "Prior"), alpha = 0.9) +
    geom_vline(xintercept = 0.7) + # true value of bias
    theme_bw() +
    labs(title = "w2 posterior and prior",
         x = "w2",
         y = "Density")

plt4 <- grid.arrange(plt1, plt2, plt3, nrow = 3)
plt4

# save to drive
ggsave("figures/Density_WB_real.png", plt4, width = 10, height = 10)
```

