# Assignment 3: Multilevel modelling

```{r setup}
pacman::p_load(tidyverse, cmdstanr, brms)
```


```{r data}
df <- read_csv("data/sc_df_clean.csv")

# fixing error in the data (some trials does not get feedback)
```
Some trials does not get feedback (NAs) - they are a group in themselves.
Some of the feedbacks are 0s - they needs to be fixed.

```{r}
bias <- seq(0, length(df$FirstRating), 1)
w1 <- 0.5
w2 <- 0.5

WeightedBayes_f <- function(bias, Source1, Source2, w1, w2){
    w1 <- (w1 - 0.5)*2
    w2 <- (w2 - 0.5)*2
    # Define or load the inv_logit_scaled and logit_scaled functions here
    outcome <- inv_logit_scaled(bias + w1 * logit_scaled(Source1) + w2 * logit_scaled(Source2)) 
    return(outcome)
}

for (i in seq(nrow(df))) {
    outcome <- WeightedBayes_f(bias, df$FirstRating, df$SecondRating, w1, w2)
    df$sim_outsome <- outcome
}

for 


view(df)
```

```{r}

outcome <- WeightedBayes_f(bias, df$FirstRating, df$SecondRating, w1, w2)

df$sim_outsome <- outcome

```

```{r}

simpleBayes_f <- function(bias, Source1, Source2){
    outcome <- inv_logit_scaled(bias + logit_scaled(Source1) + logit_scaled(Source2)) 
    return(outcome)
}



bias <- 0
Source1 <- seq(0.1,0.9, 0.1) 
Source2<- seq(0.1,0.9, 0.1) 
w1 <- seq(0.5, 1, 0.1)
w2 <- seq(0.5, 1, 0.1)

db <- expand.grid(bias = bias, Source1 = Source1, Source2 = Source2, w1 = w1, w2 = w2)

for (n in seq(nrow(db))) {
    db$belief[n] <- WeightedBayes_f(db$bias [n], db$Source1[n], db$Source2 [n], db$w1[n], db$w2 [n]) 
    db$binary[n] <- rbinom(1,1, db$belief [n])
}


```

```{r}

```

Possible effect and limitations of the model:
If you rate someone as being 5, and group does the same, you are more likely to rate the person as 5 or higher the next time, because a majority of people rate x person as being above average trustworthy.
The choices of trustworthiness are not independent of each other, but are influenced by the group rating. 

















```{r}
trials <- 150

for i in seq(trials) {
    source1 = rnorm(1, 0.5, 0.1)
}

plot(rnorm(1, 0.5, 0.1))
```

```{r}
```